Terrain is a stronger predictor of peat depth than airborne radiometrics in Norwegian landscapes

### 2025-05-27 ms/ms.tex
address RC2.3: add configurations and configuration comparisons

%
%% \usepackage commands included in the copernicus.cls:
%\usepackage[german, english]{babel}
%\usepackage{tabularx}
%\usepackage{cancel}
%\usepackage{multirow}
%\usepackage{supertabular}
%\usepackage{algorithmic}
%\usepackage{algorithm}
%\usepackage{amsthm}
%\usepackage{float}
%\usepackage{subfig}
%\usepackage{rotating}

@@

For both sites, we compared the performance of models with five different configurations of predictors.
Specifically, we trained models with:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \emph{DMK} (2 variables)
\item
  \emph{terrain} (21 variables)
\item
  \emph{terrain + DMK} (23 variables)
\item
  \emph{terrain + radiometric} (25 variables)
\item
  \emph{all predictors} (27 variables)
\end{enumerate}

These different configurations simulate different scenarios of data availability.
Comparing the different configurations allowed us to isolate the added value of each of the predictor groups.
We did not explore configurations comprising radiometrics without terrain, because LiDAR terrain surveys typically precede airborne radiometric surveys.
The models with only DMK peat depth class were simple linear models rather than RFs, and served to provide a fair comparison between the accuracy of the RF models and the existing national map of peat depth, calibrated on the same data.

We used a spatial cross-validation scheme to evaluate model performance \citep{wadouxSpatialCrossvalidationNot2021, meyerMachineLearningbasedGlobal2022}.
To set the folds we used k-Means Nearest Neighbor Distance Matching (kNNDM), which aims to mimic the spatial prediction task that is defined as the goal \citep{linnenbrinkKNNDMCVKfold2024}.
In particular, kNNDM looks for the spatial assignment of training data to folds that minimizes the difference between two distributions: nearest neighbor distances between training and test locations in the cross-validation, and nearest neighbor distances between training and prediction locations for the model.
That way, the spatial separation between folds is similar to the separation between training and prediction locations -- which increases the quality of the map accuracy estimate \citep{linnenbrinkKNNDMCVKfold2024}.
For spatially clustered training data, this approach strikes a balance between the risk of optimistic metrics from random cross-validation and the risk of pessimistic metrics from other forms of spatial cross-validation \citep{wadouxSpatialCrossvalidationNot2021}.
We implemented the kNNDM with the \emph{CAST} R package \citep[v1.0,][]{meyerCASTPackageTraining2024}, setting prediction locations to all AR5 mire cells in the study area, and choosing a number of folds (5--20) that produced the best match between the two NND distributions.
From the cross-validation we quantified mean absolute error (error magnitude, original scale), R\textsuperscript{2} (explained variation, standardized scale), and Lin's concordance correlation coefficient (error magnitude and explained variation, standardized scale).

@@

For Skrimfjella, the best predictor configuration was \emph{all predictors}, followed closely by \emph{terrain + radiometric}.
The performance gap between the \emph{terrain + DMK} configuration and the \emph{terrain} configuration was similarly small.
Compared to the \emph{terrain} configuration, the \emph{terrain + radiometric} configuration improved concordance correlation by 0.04, R\textsuperscript{2} by 0.04, and mean absolute error by 1 cm.
\emph{DMK} was a very poor predictor of peat depth even though it was calibrated to measured depths, with a concordance correlation coefficient of 0.0077.

For Ørskogfjellet, the best predictor configuration was \emph{terrain + DMK}, followed by \emph{terrain}.
Adding radiometric predictors to these configurations worsened model performance, especially in terms of concordance correlation and R\textsuperscript{2}.
By itself, \emph{DMK class} produced a concordance correlation coefficient of 0.17 (compared to 0.008 at Skrimfjella), but the worst mean absolute error of any model at either site: 77 cm.

@@

Ground-based measurements also have uncertainty \citep{parryEvaluatingApproachesEstimating2014}, but our data show that this error was smaller than the variation in peat depth over short distances (e.g.~100 m).
Independent probing and GPR measurements up to 2 m apart were differed by an average of 29--42 cm (Fig. \ref{fig:GPRwavevelocity}).
By comparison, the average variation among measurements 100 m apart was 70--106 cm (Fig. \ref{fig:semivariograms}).
We also note that cell-level measurement error that is correlated with predictors (e.g.~underestimated depth at high Multi-Resolution Valley Bottom Flatness), will go undetected by cross-validation and does not account for the limited accuracy of our models.
Ultimately, the effect of measurement error falls mostly outside the scope of DSM, since it can only be evaluated with higher-quality, independent data.

@@

Radiometric data had no predictive value at Ørskogfjellet, while at Skrimfjella they had minor influence in a relatively weak model.
The bedrock is more homogeneous at Ørskogfjellet than at Skrimfjella, so uneven radiogenesis is not a viable explanation for the differences between sites nor the poor performance in general \citep{beamishEnvironmentalRadioactivityUK2014, reinhardtGammaraySpectrometryVersatile2019}.
To the degree that radiometrics had predictive value at Skrimfjella, it appears that they were most valuable near the extremes of the depth distribution, since their inclusion improved R\textsuperscript{2} more than mean absolute error.
All four variables were highly correlated within the peatland parts of our study sites, so there could be no large differences in their predictive value.
This contrasts with Koganti et al. \citeyearpar{kogantiMappingPeatDepth2023}, who found that radiometric total count was a much better predictor than potassium ground concentration.

@@

At both our sites, LiDAR-derived terrain variables predicted peat depth much better than radiometric variables (Fig. \ref{fig:varImp}).
Elevation was the most important predictor at both sites, and peat depth showed non-monotonic responses to changes in elevation (Fig. \ref{fig:pdps}).
We believe that the idiosyncratic elevational relationships we detected are mostly not generalizable beyond the study areas, because we see no simple mechanism (e.g.~via climate) to explain the observed patterns.
For example, the increase in peat depth from 350 to 700 m.a.s.l. at Skrimfjella is opposite to the general pattern of deeper peats in lowland than upland Norway \citep{lyngstadBeskrivelserAvTorvmassivenheter2023}.
Moreover, elevation at Ørskogfjellet seems to interact with other variables (non-parallel individual conditional expectation lines), complicating its interpretation.
Relationships between elevation and peat depth have previously shown opposite shapes in different areas \citep{finlaysonEstimatingOrganicSurface2021}.
Nonetheless, a relationship that is not generalizable beyond the mapping area is still useful for DSM, as long as it is evaluated to demonstrate its robustness for the predictive task (e.g.~through kNNDM spatial cross-validation).

@@

Finally, we briefly want to highlight the need for peatland extent mapping and peat depth mapping to be better integrated.
Since peatland extent is defined by non-zero peat depth \citep[the specific threshold varies by definition,][]{minasnyMappingMonitoringPeatland2024}, the lateral and vertical dimensions are fundamentally linked.
The goal, therefore, should be a unified prediction framework for extent and depth.
The distribution of peat depths across full landscapes is zero-inflated, and research is needed to determine whether it is more efficient to parameterize a single model of peat depth (with a larger, generalized dataset) or to break down the problem into a hurdle model by classifying zero depth and then regressing non-zero depths (with smaller, specialized datasets).
We caution against reducing continuous depth predictions to arbitrary classes \citep[as in][]{ivanovsModelingGeospatialDistribution2024, karjalainenComparisonTwoGammaray2025}, since classes can be derived from continuous predictions.

\codedataavailability{R code used in this study is available at \url{https://github.com/julienvollering/DSMdepth}. The depth measurements we used are archived at \doi{10.6073/pasta/6ce440152f693f2156bf5b692a2e7917} and follow data and metadata standards.} %% use this section when having data sets and software code available

>>>>

%%\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
%
%% \usepackage commands included in the copernicus.cls:
%\usepackage[german, english]{babel}
%\usepackage{tabularx}
%\usepackage{cancel}
%\usepackage{multirow}
%\usepackage{supertabular}
%\usepackage{algorithmic}
%\usepackage{algorithm}
%\usepackage{amsthm}
%\usepackage{float}
%\usepackage{subfig}
%\usepackage{rotating}

@@

For both sites, we compared the performance of models with different predictor configurations, where each configuration was one of the seven combinations of the three predictor groups: (1) radiometric (2) terrain and (3) DMK.
Comparing the different configurations allowed us to isolate the added value of each of the predictor groups.
The models with only DMK peat depth class were simple linear models rather than RFs, and served to provide a fair comparison between the accuracy of the RF models and the existing national map of peat depth, calibrated on the same data.

We used a spatial cross-validation scheme to evaluate model performance \citep{wadouxSpatialCrossvalidationNot2021, meyerMachineLearningbasedGlobal2022}.
To set the folds we used k-Means Nearest Neighbor Distance Matching (kNNDM), which aims to mimic the spatial prediction task that is defined as the goal \citep{linnenbrinkKNNDMCVKfold2024}.
In particular, kNNDM looks for the spatial assignment of training data to folds that minimizes the difference between two distributions: nearest neighbor distances between training and test locations in the cross-validation, and nearest neighbor distances between training and prediction locations for the model.
That way, the spatial separation between folds is similar to the separation between training and prediction locations -- which increases the quality of the map accuracy estimate \citep{linnenbrinkKNNDMCVKfold2024}.
For spatially clustered training data, this approach strikes a balance between the risk of optimistic metrics from random cross-validation and the risk of pessimistic metrics from other forms of spatial cross-validation \citep{wadouxSpatialCrossvalidationNot2021}.
We implemented the kNNDM with the \emph{CAST} R package \citep[v1.0,][]{meyerCASTPackageTraining2024}, setting prediction locations to all AR5 mire cells in the study area, and choosing a number of folds (5--20) that produced the best match between the two NND distributions.
From the cross-validation we quantified mean absolute error (error magnitude, original scale), R\textsuperscript{2} (explained variation, standardized scale), and Lin's concordance correlation coefficient (error magnitude and explained variation, standardized scale).
We formally assessed the effect of predictor configuration on performance metrics using mixed-effects models to account for the cross-validation fold structure, and testing pairwise differences between configurations.

@@

For Skrimfjella, the best predictor configuration was \emph{all predictors}, followed closely by \emph{terrain + radiometric} (Fig. \ref{fig:modelMetrics}).
The \emph{terrain} configuration outperformed the \emph{radiometric} configuration by 0.27 in concordance correlation coefficient (\(p < 0.001\); Appendix Table \ref{tab:pairwiseSkrimCCC}), by 0.16 in R\textsuperscript{2} (\(p = 0.190\); Appendix Table \ref{tab:pairwiseSkrimRsq}), and by 10 cm in mean absolute error (\(p = 0.038\); Appendix Table \ref{tab:pairwiseSkrimMAE}).
The \emph{terrainDMK} configuration outperformed the \emph{radiometricDMK} configuration by 0.32 in concordance correlation coefficient (\(p < 0.001\); Appendix Table \ref{tab:pairwiseSkrimCCC}), by 0.11 in R\textsuperscript{2} (\(p = 0.605\); Appendix Table \ref{tab:pairwiseSkrimRsq}), and by 11 cm in mean absolute error (\(p = 0.011\); Appendix Table \ref{tab:pairwiseSkrimMAE}).

For Ørskogfjellet, the best predictor configuration was \emph{terrain + DMK}, followed by \emph{terrain} (Fig. \ref{fig:modelMetrics}).
Adding radiometric predictors to these configurations slightly worsened model performance.
The \emph{terrain} configuration outperformed the \emph{radiometric} configuration by 0.24 in concordance correlation coefficient (\(p = 0.068\); Appendix Table \ref{tab:pairwiseOrskogCCC}), by 0.24 in R\textsuperscript{2} (\(p = 0.004\); Appendix Table \ref{tab:pairwiseOrskogRsq}), and by 25 cm in mean absolute error (\(p = 0.086\); Appendix Table \ref{tab:pairwiseOrskogMAE}).
The \emph{terrainDMK} configuration outperformed the \emph{radiometricDMK} configuration by 0.28 in concordance correlation coefficient (\(p = 0.015\); Appendix Table \ref{tab:pairwiseOrskogCCC}), by 0.27 in R\textsuperscript{2} (\(p < 0.001\); Appendix Table \ref{tab:pairwiseOrskogRsq}), and by 28 cm in mean absolute error (\(p = 0.031\); Appendix Table \ref{tab:pairwiseOrskogMAE}).

@@

Ground-based measurements also have uncertainty \citep{parryEvaluatingApproachesEstimating2014}, but our data show that this error was smaller than the variation in peat depth over short distances (e.g.~100 m).
Independent probing and GPR measurements up to 2 m apart differed by an average of 29--42 cm (Fig. \ref{fig:GPRwavevelocity}).
By comparison, the average variation among measurements 100 m apart was 70--106 cm (Fig. \ref{fig:semivariograms}).
We also note that cell-level measurement error that is correlated with predictors (e.g.~underestimated depth at high Multi-Resolution Valley Bottom Flatness), will go undetected by cross-validation and does not account for the limited accuracy of our models.
Ultimately, the effect of measurement error falls mostly outside the scope of DSM, since it can only be evaluated with higher-quality, independent data.

@@

Radiometric data had little to no predictive value at either site (poor performance of \emph{radiometric} configuration), although they did contribute to the best model at Skrimfjella (marginal improvement in \emph{all predictors} configuration compared to \emph{terrain + DMK}).
These results contrast with earlier studies that found that radiometrics were useful predictors of peat depth \citep{keaneySpatialStatisticsEstimate2013, gatisMappingUplandPeat2019, kogantiMappingPeatDepth2023}.
The bedrock is more homogeneous at Ørskogfjellet than at Skrimfjella, so uneven radiogenesis is not a viable explanation for the differences between sites nor the poor performance in general \citep{beamishEnvironmentalRadioactivityUK2014, reinhardtGammaraySpectrometryVersatile2019}.
All four radiometric variables were highly correlated within the peatland parts of our study sites, so there could be no large differences in their predictive value.
This contrasts with Koganti et al. \citeyearpar{kogantiMappingPeatDepth2023}, who found that radiometric total count was a much better predictor than potassium ground concentration.

@@

At both our sites, LiDAR-derived terrain variables predicted peat depth much better than radiometric variables (Fig. \ref{fig:modelMetrics}).
Elevation was the most important predictor at both sites, and peat depth showed non-monotonic responses to changes in elevation (Figs. \ref{fig:varImp}-\ref{fig:pdps}).
We believe that the idiosyncratic elevational relationships we detected are mostly not generalizable beyond the study areas, because we see no simple mechanism (e.g.~via climate) to explain the observed patterns.
For example, the increase in peat depth from 350 to 700 m.a.s.l. at Skrimfjella is opposite to the general pattern of deeper peats in lowland than upland Norway \citep{lyngstadBeskrivelserAvTorvmassivenheter2023}.
Moreover, elevation at Ørskogfjellet seems to interact with other variables (non-parallel individual conditional expectation lines), complicating its interpretation.
Relationships between elevation and peat depth have previously shown opposite shapes in different areas \citep{finlaysonEstimatingOrganicSurface2021}.
Nonetheless, a relationship that is not generalizable beyond the mapping area is still useful for DSM, as long as it is evaluated to demonstrate its robustness for the predictive task (e.g.~through kNNDM spatial cross-validation).

@@

Finally, we briefly want to highlight the need for peatland extent mapping and peat depth mapping to be better integrated.
Since peatland extent is defined by non-zero peat depth \citep[the specific threshold varies by definition,][]{minasnyMappingMonitoringPeatland2024}, the lateral and vertical dimensions are fundamentally linked.
The goal, therefore, should be a unified prediction framework for extent and depth.
Research is needed to determine whether it is better to parameterize a single model of peat depth across a full landscape, or to break down the problem into a hurdle model by classifying zero depth and then regressing non-zero depths.
Though peatland definitions may encourage reducing continuous depth predictions to arbitrary classes, we caution against this practice \citep[as in][]{ivanovsModelingGeospatialDistribution2024, karjalainenComparisonTwoGammaray2025}.

\newpage
\appendix
\section{}

\appendixtables
\begin{table}[h]
\caption{Attributes of the radiometric surveys, as reported in Baranwal et al., \citeyearpar{baranwalHelicopterborneMagneticElectromagnetic2013} and Ofstad \citeyearpar{ofstadHelicopterborneMagneticRadiometric2015}.}
\begin{tabular}{lll}
\hline
& Skrimfjella & Ørskogfjellet              \\ \cline{2-3} 
Survey period                            & 2008-2011   & December 2014–January 2015 \\
Average flight altitude (m)              & 75          & 80                         \\
Average flight speed (\unit{km\,h^{-1}}) & 108         & 88                         \\
Flight line spacing (m)                  & 200         & 200                        \\ \hline
\end{tabular}
\label{tab:radSurveys}
\end{table}
\clearpage

\appendixfigures
\begin{figure}[ht]
\centering
\includegraphics[width=0.8\textwidth]{figures/GPRwavevelocity.pdf}
\caption{\label{fig:GPRwavevelocity}Calibration of GPR wave velocity at Skrimfjella (a) and Ørskogfjellet (b) by regression of probed depth against wave travel time.}
\end{figure}
\clearpage

\newpage
\section{}

\begin{table}[h]

\caption{\label{tab:pairwiseSkrimCCC}Pairwise comparisons of the concordance correlation coefficient among models for Skrimfjella, based on 10-fold cross-validation. The comparisons use estimated marginal means with Tukey HSD correction and Kenward-Roger degrees of freedom approximation (\texttt{emmeans} package) on mixed-effects models (\texttt{lme4} package) to account for the cross-validation fold structure.}
\centering
\begin{tabular}[t]{lrrrrll}
\hline
Comparison & Estimate & Std.error & Df & Statistic & Adj.p.value & Significance\\
\hline
DMK - Radiometric & 0.02 & 0.05 & 114 & 0.44 & 0.999 & \\
DMK - RadiometricDMK & 0.06 & 0.05 & 114 & 1.19 & 0.896 & \\
DMK - RadiometricTerrain & -0.29 & 0.05 & 114 & -5.84 & < 0.001 & ***\\
DMK - RadiometricTerrainDMK & -0.29 & 0.05 & 114 & -5.81 & < 0.001 & ***\\
DMK - Terrain & -0.25 & 0.05 & 114 & -5.03 & < 0.001 & ***\\
DMK - TerrainDMK & -0.26 & 0.05 & 114 & -5.17 & < 0.001 & ***\\
Radiometric - RadiometricDMK & 0.04 & 0.05 & 114 & 0.75 & 0.989 & \\
Radiometric - RadiometricTerrain & -0.31 & 0.05 & 114 & -6.28 & < 0.001 & ***\\
Radiometric - RadiometricTerrainDMK & -0.31 & 0.05 & 114 & -6.25 & < 0.001 & ***\\
Radiometric - Terrain & -0.27 & 0.05 & 114 & -5.47 & < 0.001 & ***\\
Radiometric - TerrainDMK & -0.28 & 0.05 & 114 & -5.62 & < 0.001 & ***\\
RadiometricDMK - RadiometricTerrain & -0.35 & 0.05 & 114 & -7.03 & < 0.001 & ***\\
RadiometricDMK - RadiometricTerrainDMK & -0.35 & 0.05 & 114 & -7.00 & < 0.001 & ***\\
RadiometricDMK - Terrain & -0.31 & 0.05 & 114 & -6.22 & < 0.001 & ***\\
RadiometricDMK - TerrainDMK & -0.32 & 0.05 & 114 & -6.37 & < 0.001 & ***\\
RadiometricTerrain - RadiometricTerrainDMK & 0.00 & 0.05 & 114 & 0.03 & 1 & \\
RadiometricTerrain - Terrain & 0.04 & 0.05 & 114 & 0.81 & 0.983 & \\
RadiometricTerrain - TerrainDMK & 0.03 & 0.05 & 114 & 0.67 & 0.994 & \\
RadiometricTerrainDMK - Terrain & 0.04 & 0.05 & 114 & 0.79 & 0.986 & \\
RadiometricTerrainDMK - TerrainDMK & 0.03 & 0.05 & 114 & 0.64 & 0.995 & \\
Terrain - TerrainDMK & -0.01 & 0.05 & 114 & -0.15 & 1 & \\
\hline
\end{tabular}
\end{table}
\clearpage

\begin{table}[h]

\caption{\label{tab:pairwiseSkrimRsq}Pairwise comparisons of $R^2$ among models for Skrimfjella, based on 10-fold cross-validation. The comparisons use estimated marginal means with Tukey HSD correction and Kenward-Roger degrees of freedom approximation (\texttt{emmeans} package) on mixed-effects models (\texttt{lme4} package) to account for the cross-validation fold structure.}
\centering
\begin{tabular}[t]{lrrrrrl}
\hline
Comparison & Estimate & Std.error & Df & Statistic & Adj.p.value & Significance\\
\hline
DMK - Radiometric & 0.04 & 0.08 & 108.23 & 0.55 & 0.998 & \\
DMK - RadiometricDMK & 0.01 & 0.08 & 108.23 & 0.08 & 1.000 & \\
DMK - RadiometricTerrain & -0.16 & 0.08 & 108.23 & -2.07 & 0.378 & \\
DMK - RadiometricTerrainDMK & -0.16 & 0.08 & 108.23 & -2.11 & 0.357 & \\
DMK - Terrain & -0.11 & 0.08 & 108.23 & -1.48 & 0.755 & \\
DMK - TerrainDMK & -0.10 & 0.08 & 108.23 & -1.34 & 0.830 & \\
Radiometric - RadiometricDMK & -0.04 & 0.06 & 105.01 & -0.56 & 0.998 & \\
Radiometric - RadiometricTerrain & -0.20 & 0.06 & 105.01 & -3.16 & 0.033 & *\\
Radiometric - RadiometricTerrainDMK & -0.20 & 0.06 & 105.01 & -3.20 & 0.029 & *\\
Radiometric - Terrain & -0.16 & 0.06 & 105.01 & -2.45 & 0.190 & \\
Radiometric - TerrainDMK & -0.15 & 0.06 & 105.01 & -2.28 & 0.263 & \\
RadiometricDMK - RadiometricTerrain & -0.17 & 0.06 & 105.01 & -2.60 & 0.137 & \\
RadiometricDMK - RadiometricTerrainDMK & -0.17 & 0.06 & 105.01 & -2.64 & 0.124 & \\
RadiometricDMK - Terrain & -0.12 & 0.06 & 105.01 & -1.89 & 0.494 & \\
RadiometricDMK - TerrainDMK & -0.11 & 0.06 & 105.01 & -1.72 & 0.605 & \\
RadiometricTerrain - RadiometricTerrainDMK & 0.00 & 0.06 & 105.01 & -0.04 & 1.000 & \\
RadiometricTerrain - Terrain & 0.05 & 0.06 & 105.01 & 0.71 & 0.992 & \\
RadiometricTerrain - TerrainDMK & 0.06 & 0.06 & 105.01 & 0.88 & 0.975 & \\
RadiometricTerrainDMK - Terrain & 0.05 & 0.06 & 105.01 & 0.75 & 0.989 & \\
RadiometricTerrainDMK - TerrainDMK & 0.06 & 0.06 & 105.01 & 0.92 & 0.968 & \\
Terrain - TerrainDMK & 0.01 & 0.06 & 105.01 & 0.17 & 1.000 & \\
\hline
\end{tabular}
\end{table}
\clearpage

\begin{table}[h]

\caption{\label{tab:pairwiseSkrimMAE}Pairwise comparisons of mean absolute error among models for Skrimfjella, based on 10-fold cross-validation. The comparisons use estimated marginal means with Tukey HSD correction and Kenward-Roger degrees of freedom approximation (\texttt{emmeans} package) on mixed-effects models (\texttt{lme4} package) to account for the cross-validation fold structure.}
\centering
\begin{tabular}[t]{lrrrrrl}
\hline
Comparison & Estimate & Std.error & Df & Statistic & Adj.p.value & Significance\\
\hline
DMK - Radiometric & -2.91 & 3.13 & 114 & -0.93 & 0.967 & \\
DMK - RadiometricDMK & -3.69 & 3.13 & 114 & -1.18 & 0.902 & \\
DMK - RadiometricTerrain & 8.10 & 3.13 & 114 & 2.58 & 0.140 & \\
DMK - RadiometricTerrainDMK & 8.21 & 3.13 & 114 & 2.62 & 0.130 & \\
DMK - Terrain & 6.81 & 3.13 & 114 & 2.17 & 0.319 & \\
DMK - TerrainDMK & 7.33 & 3.13 & 114 & 2.34 & 0.235 & \\
Radiometric - RadiometricDMK & -0.77 & 3.13 & 114 & -0.25 & 1.000 & \\
Radiometric - RadiometricTerrain & 11.02 & 3.13 & 114 & 3.51 & 0.011 & *\\
Radiometric - RadiometricTerrainDMK & 11.13 & 3.13 & 114 & 3.55 & 0.010 & *\\
Radiometric - Terrain & 9.72 & 3.13 & 114 & 3.10 & 0.038 & *\\
Radiometric - TerrainDMK & 10.24 & 3.13 & 114 & 3.27 & 0.024 & *\\
RadiometricDMK - RadiometricTerrain & 11.79 & 3.13 & 114 & 3.76 & 0.005 & **\\
RadiometricDMK - RadiometricTerrainDMK & 11.90 & 3.13 & 114 & 3.80 & 0.004 & **\\
RadiometricDMK - Terrain & 10.50 & 3.13 & 114 & 3.35 & 0.018 & *\\
RadiometricDMK - TerrainDMK & 11.01 & 3.13 & 114 & 3.51 & 0.011 & *\\
RadiometricTerrain - RadiometricTerrainDMK & 0.11 & 3.13 & 114 & 0.04 & 1.000 & \\
RadiometricTerrain - Terrain & -1.29 & 3.13 & 114 & -0.41 & 1.000 & \\
RadiometricTerrain - TerrainDMK & -0.77 & 3.13 & 114 & -0.25 & 1.000 & \\
RadiometricTerrainDMK - Terrain & -1.40 & 3.13 & 114 & -0.45 & 0.999 & \\
RadiometricTerrainDMK - TerrainDMK & -0.88 & 3.13 & 114 & -0.28 & 1.000 & \\
Terrain - TerrainDMK & 0.52 & 3.13 & 114 & 0.17 & 1.000 & \\
\hline
\end{tabular}
\end{table}
\clearpage

\begin{table}[h]

\caption{\label{tab:pairwiseOrskogCCC}Pairwise comparisons of the concordance correlation coefficient among models for Ørskogfjellet, based on 10-fold cross-validation. The comparisons use estimated marginal means with Tukey HSD correction and Kenward-Roger degrees of freedom approximation (\texttt{emmeans} package) on mixed-effects models (\texttt{lme4} package) to account for the cross-validation fold structure.}
\centering
\begin{tabular}[t]{lrrrrrl}
\hline
Comparison & Estimate & Std.error & Df & Statistic & Adj.p.value & Significance\\
\hline
DMK - Radiometric & 0.05 & 0.08 & 54 & 0.61 & 0.996 & \\
DMK - RadiometricDMK & 0.07 & 0.08 & 54 & 0.84 & 0.979 & \\
DMK - RadiometricTerrain & -0.14 & 0.08 & 54 & -1.79 & 0.559 & \\
DMK - RadiometricTerrainDMK & -0.16 & 0.08 & 54 & -1.95 & 0.458 & \\
DMK - Terrain & -0.19 & 0.08 & 54 & -2.33 & 0.248 & \\
DMK - TerrainDMK & -0.22 & 0.08 & 54 & -2.67 & 0.127 & \\
Radiometric - RadiometricDMK & 0.02 & 0.08 & 54 & 0.24 & 1.000 & \\
Radiometric - RadiometricTerrain & -0.19 & 0.08 & 54 & -2.40 & 0.220 & \\
Radiometric - RadiometricTerrainDMK & -0.21 & 0.08 & 54 & -2.55 & 0.161 & \\
Radiometric - Terrain & -0.24 & 0.08 & 54 & -2.94 & 0.068 & \\
Radiometric - TerrainDMK & -0.26 & 0.08 & 54 & -3.27 & 0.029 & *\\
RadiometricDMK - RadiometricTerrain & -0.21 & 0.08 & 54 & -2.63 & 0.136 & \\
RadiometricDMK - RadiometricTerrainDMK & -0.23 & 0.08 & 54 & -2.79 & 0.096 & \\
RadiometricDMK - Terrain & -0.26 & 0.08 & 54 & -3.17 & 0.037 & *\\
RadiometricDMK - TerrainDMK & -0.28 & 0.08 & 54 & -3.51 & 0.015 & *\\
RadiometricTerrain - RadiometricTerrainDMK & -0.01 & 0.08 & 54 & -0.16 & 1.000 & \\
RadiometricTerrain - Terrain & -0.04 & 0.08 & 54 & -0.54 & 0.998 & \\
RadiometricTerrain - TerrainDMK & -0.07 & 0.08 & 54 & -0.88 & 0.975 & \\
RadiometricTerrainDMK - Terrain & -0.03 & 0.08 & 54 & -0.38 & 1.000 & \\
RadiometricTerrainDMK - TerrainDMK & -0.06 & 0.08 & 54 & -0.72 & 0.991 & \\
Terrain - TerrainDMK & -0.03 & 0.08 & 54 & -0.33 & 1.000 & \\
\hline
\end{tabular}
\end{table}
\clearpage

\begin{table}[h]

\caption{\label{tab:pairwiseOrskogRsq}Pairwise comparisons of $R^2$ among models for Ørskogfjellet, based on 10-fold cross-validation. The comparisons use estimated marginal means with Tukey HSD correction and Kenward-Roger degrees of freedom approximation (\texttt{emmeans} package) on mixed-effects models (\texttt{lme4} package) to account for the cross-validation fold structure.}
\centering
\begin{tabular}[t]{lrrrrll}
\hline
Comparison & Estimate & Std.error & Df & Statistic & Adj.p.value & Significance\\
\hline
DMK - Radiometric & 0.08 & 0.06 & 54 & 1.36 & 0.818 & \\
DMK - RadiometricDMK & 0.08 & 0.06 & 54 & 1.35 & 0.823 & \\
DMK - RadiometricTerrain & -0.10 & 0.06 & 54 & -1.72 & 0.607 & \\
DMK - RadiometricTerrainDMK & -0.11 & 0.06 & 54 & -1.91 & 0.485 & \\
DMK - Terrain & -0.15 & 0.06 & 54 & -2.62 & 0.14 & \\
DMK - TerrainDMK & -0.19 & 0.06 & 54 & -3.29 & 0.028 & *\\
Radiometric - RadiometricDMK & 0.00 & 0.06 & 54 & -0.01 & 1 & \\
Radiometric - RadiometricTerrain & -0.18 & 0.06 & 54 & -3.08 & 0.048 & *\\
Radiometric - RadiometricTerrainDMK & -0.19 & 0.06 & 54 & -3.27 & 0.029 & *\\
Radiometric - Terrain & -0.24 & 0.06 & 54 & -3.98 & 0.004 & **\\
Radiometric - TerrainDMK & -0.28 & 0.06 & 54 & -4.65 & < 0.001 & ***\\
RadiometricDMK - RadiometricTerrain & -0.18 & 0.06 & 54 & -3.07 & 0.049 & *\\
RadiometricDMK - RadiometricTerrainDMK & -0.19 & 0.06 & 54 & -3.26 & 0.03 & *\\
RadiometricDMK - Terrain & -0.24 & 0.06 & 54 & -3.97 & 0.004 & **\\
RadiometricDMK - TerrainDMK & -0.27 & 0.06 & 54 & -4.64 & < 0.001 & ***\\
RadiometricTerrain - RadiometricTerrainDMK & -0.01 & 0.06 & 54 & -0.19 & 1 & \\
RadiometricTerrain - Terrain & -0.05 & 0.06 & 54 & -0.90 & 0.971 & \\
RadiometricTerrain - TerrainDMK & -0.09 & 0.06 & 54 & -1.57 & 0.702 & \\
RadiometricTerrainDMK - Terrain & -0.04 & 0.06 & 54 & -0.71 & 0.991 & \\
RadiometricTerrainDMK - TerrainDMK & -0.08 & 0.06 & 54 & -1.38 & 0.809 & \\
Terrain - TerrainDMK & -0.04 & 0.06 & 54 & -0.67 & 0.994 & \\
\hline
\end{tabular}
\end{table}
\clearpage

\begin{table}[h]

\caption{\label{tab:pairwiseOrskogMAE}Pairwise comparisons of mean absolute error among models for Ørskogfjellet, based on 10-fold cross-validation. The comparisons use estimated marginal means with Tukey HSD correction and Kenward-Roger degrees of freedom approximation (\texttt{emmeans} package) on mixed-effects models (\texttt{lme4} package) to account for the cross-validation fold structure.}
\centering
\begin{tabular}[t]{lrrrrrl}
\hline
Comparison & Estimate & Std.error & Df & Statistic & Adj.p.value & Significance\\
\hline
DMK - Radiometric & -6.93 & 8.69 & 54 & -0.80 & 0.984 & \\
DMK - RadiometricDMK & -7.63 & 8.69 & 54 & -0.88 & 0.974 & \\
DMK - RadiometricTerrain & 14.31 & 8.69 & 54 & 1.65 & 0.653 & \\
DMK - RadiometricTerrainDMK & 15.83 & 8.69 & 54 & 1.82 & 0.539 & \\
DMK - Terrain & 17.71 & 8.69 & 54 & 2.04 & 0.403 & \\
DMK - TerrainDMK & 20.55 & 8.69 & 54 & 2.37 & 0.233 & \\
Radiometric - RadiometricDMK & -0.70 & 8.69 & 54 & -0.08 & 1.000 & \\
Radiometric - RadiometricTerrain & 21.24 & 8.69 & 54 & 2.45 & 0.200 & \\
Radiometric - RadiometricTerrainDMK & 22.76 & 8.69 & 54 & 2.62 & 0.140 & \\
Radiometric - Terrain & 24.64 & 8.69 & 54 & 2.84 & 0.086 & \\
Radiometric - TerrainDMK & 27.48 & 8.69 & 54 & 3.16 & 0.039 & *\\
RadiometricDMK - RadiometricTerrain & 21.95 & 8.69 & 54 & 2.53 & 0.170 & \\
RadiometricDMK - RadiometricTerrainDMK & 23.47 & 8.69 & 54 & 2.70 & 0.118 & \\
RadiometricDMK - Terrain & 25.35 & 8.69 & 54 & 2.92 & 0.071 & \\
RadiometricDMK - TerrainDMK & 28.19 & 8.69 & 54 & 3.24 & 0.031 & *\\
RadiometricTerrain - RadiometricTerrainDMK & 1.52 & 8.69 & 54 & 0.17 & 1.000 & \\
RadiometricTerrain - Terrain & 3.40 & 8.69 & 54 & 0.39 & 1.000 & \\
RadiometricTerrain - TerrainDMK & 6.24 & 8.69 & 54 & 0.72 & 0.991 & \\
RadiometricTerrainDMK - Terrain & 1.88 & 8.69 & 54 & 0.22 & 1.000 & \\
RadiometricTerrainDMK - TerrainDMK & 4.72 & 8.69 & 54 & 0.54 & 0.998 & \\
Terrain - TerrainDMK & 2.84 & 8.69 & 54 & 0.33 & 1.000 & \\
\hline
\end{tabular}
\end{table}
\clearpage

\begin{figure}[ht]
\centering
\includegraphics{figures/semivariograms.pdf}
\caption{\label{fig:semivariograms}Empirical semivariograms of peat depth point measurements at Skrimfjella (a) and Ørskogfjellet (b).}
\end{figure}
\clearpage

\codedataavailability{\emph{Code and data availability.} R code used in this study is available at \url{https://github.com/julienvollering/DSMdepth}. The depth measurements we used are archived at \doi{10.6073/pasta/6ce440152f693f2156bf5b692a2e7917} and follow data and metadata standards.} %% use this section when having data sets and software code available
